---
title: "exp1and2"
author: "Tovly Deutsch"
date: "4/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Attribution
Portions of this code were written by the authors of Michael Becker, Andrew Nevins, and Jonathan Levine (2012) Asymmetries in generalizing alternations to and from initial syllables. Language 88:2, pp. 231–268.
That orginal code and data (not included in this repo) can be found at http://becker.phonologist.org/projects/english/.

## Setup
```{r}
exp1 = read.csv("../corpora/exp1.csv")
exp2 = read.csv("../corpora/exp2.csv", encoding = 'UTF-8')

exp1$shape_group = factor(exp1$shape_group, levels=c("P","_P","P_"))
exp2$shape = factor(exp2$shape, levels=c("mono","iamb","trochee"))
```

## CMU-ipa dictionary lookup
The CMU-ipa dict is by Alexander Piperski and can be found at https://github.com/menelik3/cmudict-ipa.
```{r}
library('utf8')
library('utils')
cmu_dict = read.csv('../corpora/cmudict-0.7b-ipa.tsv', sep = '\t', header = FALSE,
                    quote = "", encoding = 'UTF-8', colClasses = "character")
cmu_dict
cmu_lookup <- cmu_dict$V2
names(cmu_lookup) <- cmu_dict$V1
stress = TRUE
get_phonetic_form = function(word) {
  upper_word = toupper(word)
  if (upper_word %in% names(cmu_lookup)) {
    result = cmu_lookup[[upper_word]]
    # remove stress and length markers (\u02C8 is the primary stress marker)
    if (stress) {
      result = str_remove_all(result, "ˌ|ː")
    }
    else {
      result = str_remove_all(result, "\u02C8|ˌ|ː")
    }
  }
  else {
    cat(sprintf('Could not find phonetic version of "%s"\n', word))
    result = NA
  }
  return(result)
}
```
```{r}
insertStress = function(word, shape) {
  vowels = "a|e|i|ɪ|ə|ʌ|æ|ʊ|ɛ|u|ɝ|ɑ|ɔ|o"
  vowelIndices = grep(vowels, unlist(strsplit(word, split="")))
  if (length(vowelIndices) == 0) {
    stop("Error, no vowels")
  }
  # print(shape == 'iamb')
  if (shape == 'iamb' | shape == '_P') {
    lastVowel = tail(vowelIndices, n = 1)
    consonant = lastVowel - 1
    while (grepl(vowels, substr(word, consonant, consonant)) & consonant > 1) {
      consonant = consonant - 1
    }
    stressed = paste(substr(word, 1, consonant - 1), "\u02C8", substr(word, consonant, str_length(word)), sep = "")
  }
  else {
    stressed = paste("\u02C8", substr(word, 1, str_length(word)), sep = "")
  }
  return(stressed)
}
```



## write exp1 average responses (1 = voiceless, 7 = voiced) from 200 participants as tsv
```{r}

items = with(subset(exp1, type=="stim"), aggregate(rating.pl, list(item), mean));
colnames(items) = c("item", "voice")
items = merge(items, unique(exp1[,c("item","shape_group")]))
items$item = as.character(items$item)
items = items %>%
  mutate('IPA' = Vectorize(get_phonetic_form)(item))
items = items[complete.cases(items),]
# TODO fix r remaning in words with multiple prnounciations
items$IPA <- gsub("r", "ɹ", items$IPA)

write_utf8_csv <- function(df, file) {
    firstline <- paste(names(df), sep = "", collapse = "\t")
    data <- apply(df, 1, function(x) {paste(x, sep = "", collapse = "\t")})
    writeLines(c(firstline, data), file , useBytes = TRUE)
}

print(sd(items$voice))
items
# for (i in 1:nrow(items)) {
#   items[i,]$IPA = insertStress(items[i,]$IPA, items[i,]$shape_group)
# }

write_utf8_csv(items, "../corpora/stress_english_voicing_ratings.tsv")
```

## exp2 average responses (1 = voiceless, 7 = voiced) from 200 participants
```{r}
items2 = with(subset(exp2, type=="stim"), aggregate(rating1, list(item), mean));
colnames(items2) = c("item","voice")
items2 = items2[order(items2$voice),]; 
items2 = merge(items2, unique(exp2[,c("item","shape")]))
items2 = merge(items2, unique(exp2[,c("item","IPA")]))
# items2 = items2 %>% select(-shape)
items2$IPA <- gsub("ɚ", "ɝ", items2$IPA)
items2$IPA <- gsub("r", "ɹ", items2$IPA)
items2$IPA <- gsub("ʧ", "tʃ", items2$IPA)
items2$IPA <- gsub("ʤ", "dʒ", items2$IPA)
items2$shape = as.character(items2$shape)

for (i in 1:nrow(items2)) {
  items2[i,]$IPA = insertStress(items2[i,]$IPA, items2[i,]$shape)
}
# items2 = items2 %>% mutate(IPA = insertStress(IPA, shape))
# items2
# print(insertStress("floʊdəəf"))
# print(sd(items2$voice))
write_utf8_csv(items2, "../corpora/stress_wug_voicing_ratings.tsv")
```

